Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-04-06 summary.Rmd
pmlr -> jmlr

Using results from \citet{soltanolkotabi2012}, it can be easily shown that 
the subspace detection property holds for $XU$, which is an ASE of $P$. More 
specifically, if points lie exactly on mutually orthogonal subspaces, then the 
subspace detection property will hold with probability 1, and this is exactly 
the case for the PABM (Theorem \ref{theorem2}). 
Much of our work is then built on
\citeauthor{rubindelanchy2017statistical}, who describe the convergence 
behavior of the ASE of $A$ to the ASE of $P$, and \citet{pmlr-v28-wang13},
who show the necessary conditions for the subspace detection property to hold 
in noisy cases where the points lie near subspaces.

@@

\begin{algorithm}[t]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{Sparse Subspace Clustering using LASSO \cite{pmlr-v28-wang13}.}
  \KwData{Adjacency matrix $A$, number of communities $K$, 
  hyperparameter $\lambda$}
  \KwResult{Community assignments $1, ..., K$}
    Find $V$, the matrix of eigenvectors of $A$ 
    corresponding to the $K (K + 1) / 2$ most positive 
    and the $K (K - 1) / 2$ most negative eigenvalues.\;
    \For {$i = 1, ..., n$} {
      Assign $v_i^\top$ as the $i^{th}$ row of $V$. 
      Assign $V_{-i} = \begin{bmatrix} 
      v_1 & \cdots & v_{i-1} & v_{i+1} & \cdots & v_n \end{bmatrix}^\top$.\;
      Solve the LASSO problem 
      $c_i = \arg\min_{\beta} 
      \frac{1}{2} ||v_i - V_{-i} \beta||_2^2 + \lambda ||\beta||_1$.\;
      Assign $\tilde{c}_i = \begin{bmatrix} 
      c_i^{(1)} & \cdots & c_i^{(i-1)} & 0 & c_i^{(i)} & \cdots & c_i^{(n-1)}
      \end{bmatrix}^\top$ such that the superscript is the index of 
      $\tilde{c}_i$.\;
    }
    Assign 
    $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
    Compute the affinity matrix $B = \sqrt{n} (|C| + |C^\top|)$.\;
    Construct graph $G$ using $B$ as its similarity matrix.\;
    Partition $G$ into $K$ disconnected subgraphs (e.g., using edge 
    thresholding or spectral clustering).\;
    Map each partition to the community labels $1, ..., K$. 
\end{algorithm}

@@

This paper shows the connection between the PABM and the GRDPG, namely that 
a PABM graph can be represented as a union of orthogonal subspaces in an 
embedding under the GRDPG framework. We then exploited this relationship to 
develop community detection and parameter estimation methods. In fact, we can 
represent any graph with Bernoulli edges as a GRDPG, and in the PABM case, 
it turns out that this relationship leads to a very straightforward 
applications of previous work from \citeauthor{rubindelanchy2017statistical}, 
\citeauthor{soltanolkotabi2012}, and \citeauthor{pmlr-v28-wang13}, which lead 
to asymptotically correct solutions with high probability. Similar methods can 
be applied for other models, such as the Nested Block Model
\cite{noroozi2021hierarchy}.

@@

\begin{definition}[Inradius \cite{soltanolkotabi2012} \cite{pmlr-v28-wang13}]
The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents 
the inradius of the symmetric convex hull of $X$. 
\end{definition}

@@

Furthermore, Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the 
extended definition of $\mu(\cdot)$ from \citeauthor{pmlr-v28-wang13}. Then

@@

**Proof of Theorem \ref{theorem5}**. 
The basis of this proof is Theorem 6 from \citeauthor{pmlr-v28-wang13}, which 
states that the subspace detection property holds if the noise is small enough 
and the subspace inradius is greater than the subspace incoherence for each 
community $k$.

@@

Thus the conditions for the subspace detection property from Theorem 6 from
\citeauthor{pmlr-v28-wang13} are satisfied with probability 1 as $n \to \infty$.

>>>>

Using results from \citet{soltanolkotabi2012}, it can be easily shown that 
the subspace detection property holds for $XU$, which is an ASE of $P$. More 
specifically, if points lie exactly on mutually orthogonal subspaces, then the 
subspace detection property will hold with probability 1, and this is exactly 
the case for the PABM (Theorem \ref{theorem2}). 
Much of our work is then built on
\citeauthor{rubindelanchy2017statistical}, who describe the convergence 
behavior of the ASE of $A$ to the ASE of $P$, and \citet{jmlr-v28-wang13},
who show the necessary conditions for the subspace detection property to hold 
in noisy cases where the points lie near subspaces.

@@

\begin{algorithm}[t]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{Sparse Subspace Clustering using LASSO \cite{jmlr-v28-wang13}.}
  \KwData{Adjacency matrix $A$, number of communities $K$, 
  hyperparameter $\lambda$}
  \KwResult{Community assignments $1, ..., K$}
    Find $V$, the matrix of eigenvectors of $A$ 
    corresponding to the $K (K + 1) / 2$ most positive 
    and the $K (K - 1) / 2$ most negative eigenvalues.\;
    \For {$i = 1, ..., n$} {
      Assign $v_i^\top$ as the $i^{th}$ row of $V$. 
      Assign $V_{-i} = \begin{bmatrix} 
      v_1 & \cdots & v_{i-1} & v_{i+1} & \cdots & v_n \end{bmatrix}^\top$.\;
      Solve the LASSO problem 
      $c_i = \arg\min_{\beta} 
      \frac{1}{2} ||v_i - V_{-i} \beta||_2^2 + \lambda ||\beta||_1$.\;
      Assign $\tilde{c}_i = \begin{bmatrix} 
      c_i^{(1)} & \cdots & c_i^{(i-1)} & 0 & c_i^{(i)} & \cdots & c_i^{(n-1)}
      \end{bmatrix}^\top$ such that the superscript is the index of 
      $\tilde{c}_i$.\;
    }
    Assign 
    $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
    Compute the affinity matrix $B = \sqrt{n} (|C| + |C^\top|)$.\;
    Construct graph $G$ using $B$ as its similarity matrix.\;
    Partition $G$ into $K$ disconnected subgraphs (e.g., using edge 
    thresholding or spectral clustering).\;
    Map each partition to the community labels $1, ..., K$. 
\end{algorithm}

@@

This paper shows the connection between the PABM and the GRDPG, namely that 
a PABM graph can be represented as a union of orthogonal subspaces in an 
embedding under the GRDPG framework. We then exploited this relationship to 
develop community detection and parameter estimation methods. In fact, we can 
represent any graph with Bernoulli edges as a GRDPG, and in the PABM case, 
it turns out that this relationship leads to a very straightforward 
applications of previous work from \citeauthor{rubindelanchy2017statistical}, 
\citeauthor{soltanolkotabi2012}, and \citeauthor{jmlr-v28-wang13}, which lead 
to asymptotically correct solutions with high probability. Similar methods can 
be applied for other models, such as the Nested Block Model
\cite{noroozi2021hierarchy}.

@@

\begin{definition}[Inradius \cite{soltanolkotabi2012} \cite{jmlr-v28-wang13}]
The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents 
the inradius of the symmetric convex hull of $X$. 
\end{definition}

@@

Furthermore, Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the 
extended definition of $\mu(\cdot)$ from \citeauthor{jmlr-v28-wang13}. Then

@@

**Proof of Theorem \ref{theorem5}**. 
The basis of this proof is Theorem 6 from \citeauthor{jmlr-v28-wang13}, which 
states that the subspace detection property holds if the noise is small enough 
and the subspace inradius is greater than the subspace incoherence for each 
community $k$.

@@

Thus the conditions for the subspace detection property from Theorem 6 from
\citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as $n \to \infty$.
